---
title: "Probability"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 13.9

1. One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls,
and 7 yellow balls. What is the probability that the ball will be cyan?

```{r}
box <- rep(c("cyan", "magenta", "yellow"), c(3,5,7))
c <- "cyan"
mean(box %in% c)
```

2. What is the probability that the ball will not be cyan?

```{r}
box <- rep(c("cyan", "magenta", "yellow"), c(3,5,7))
others <- c("magenta", "yellow")
mean(box %in% others)
```

3. Instead of taking just one draw, consider taking two draws. You take the second draw
without returning the first draw to the box. We call this sampling without replacement.
What is the probability that the first draw is cyan and that the second draw is not cyan?

```{r}
library(gtools)
sam_space <- permutations(15,2, v=box, set=FALSE)
first <- sam_space[,1]
second <- sam_space[,2]
c <- "cyan"
others <- c("magenta", "yellow")
mean(first %in% c & second %in% others)
```

4. Now repeat the experiment, but this time, after taking the first draw and recording the
color, return it to the box and shake the box. We call this sampling with replacement. What is the probability that the first draw is cyan and that the second draw is not cyan?

```{r}
box <- rep(c("cyan", "magenta", "yellow"), c(3,5,7))
replace <- expand.grid(first_r = box, second_r = box)
first_r <- replace[,1]
second_r <- replace[,2]
c <- "cyan"
others <- c("magenta", "yellow")
mean(first_r %in% c & second_r %in% others)
```

## Exercise 13.14
1. Assume the distribution of female heights is approximated by a normal distribution with
a mean of 64 inches and a standard deviation of 3 inches. If we pick a female at random,
what is the probability that she is 5 feet or shorter?

```{r}
pnorm(5*12,64,3)
```

2. Assume the distribution of female heights is approximated by a normal distribution with
a mean of 64 inches and a standard deviation of 3 inches. If we pick a female at random,
what is the probability that she is 6 feet or taller?

```{r}
1-pnorm(6*12,64,3)
```

3. Assume the distribution of female heights is approximated by a normal distribution with
a mean of 64 inches and a standard deviation of 3 inches. If we pick a female at random,
what is the probability that she is between 61 and 67 inches?

```{r}
pnorm(67,64,3)-pnorm(61,64,3)
```

4. Repeat the exercise above, but convert everything to centimeters. That is, multiply every height, including the standard deviation, by 2.54. What is the answer now?

```{r}
pnorm(67*2.54,64*2.54,3*2.54)-pnorm(61*2.54,64*2.54,3*2.54)
```

