---
title: "Homework 8"
author: "Pakorn Sagulkoo 6380064520"
date: "11/11/2563"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise

Balls in a box
1. We had a red box containing many color balls which we don’t know how many of
them. However, we believed that picking a ball out of this box, the proportion of
times that we get a red box is 1/4. Use the Bayesian analysis to estimate the
probability of getting a red ball out of a black box if we do the experiment and gain
some knowledge as follows:

Experiment 1 performed 5 trials of picking a ball and returning it into the box,
we obtained a red ball 2 times.
Experiment 2 performed 10 trials of picking a ball and returning it into the box,
we obtained a red ball 3 times.
Experiment 3 performed 12 trials of picking a ball and returning it into the box,
we obtained a red ball 5 times.
Use the Bayesian model to estimate the probability of getting a red ball out of this
black box.

Use the Bayesian model to estimate the probability of getting a red ball out of this
black box.

  1.1. Specify your own prior distribution and posterior distributions and plot the prior distribution and all of the posterior distributions in the same graph.

```{r}
expl <- (0:1000)/1000
plot(expl, dbeta(expl,2,5),ylim=range(0:6),type="l",main="Prior and Posterior distributions")
lines(expl,dbeta(expl,4,8),col="red")
text(0.18,2.8,labels ="Prior")
text(0.3,3.3,labels="Post1",col="red")
lines(expl, dbeta(expl,7,15),col="darkgreen")
text(0.3,4.3,labels="Post2",col="darkgreen")
lines(expl, dbeta(expl,12,22),col="darkmagenta")
text(0.35,5.2,labels="Post3",col="darkmagenta")
```

  1.2. What is the estimated probability of getting a red ball out of this black box?
  
```{r}
which(dbeta(expl,12,22) == max(dbeta(expl,12,22)))/1000
```
  2. Male or female
Suppose that 40 people are randomly picked off a city street in Bangkok. Whether they are male or female is noted on 40 identical pieces of paper, put into a hat and the hat is brought to us. We have not seen these 40 people. We then take out five pieces of paper from the hat and read them - three are female.

Estimate the number of females in the original group of 40 with the Bayesian model. hypo : 50:50 (m:f)

```{r}
samprop <- seq(0, 1, length.out = 40)
plot(samprop, dbinom(x=3, prob=samprop, size=5),ylim=range(0:1),
type="l", xlab="Female prop", ylab="Density")
text(0.6,0.4,labels="likelihood")
lines(samprop, dnorm(x=samprop, mean=0.5,sd=0.1)/25,
col="blue")
text(0.5,0.2,labels="Prior",col="blue")
lh <- dbinom(x=3, prob=samprop, size=5)
pri <- dnorm(x=samprop, mean=0.5, sd=0.1)
lines(samprop, lh*pri, col="darkorange")
text(0.74,0.6,labels="Unstdard Post",col="darkorange")
unstd_post <- lh * pri
std_post <- unstd_post / sum(unstd_post)
lines(samprop, std_post, col="deeppink")
text(0.6,0.15,labels="Post",col="deeppink")
```
  The female propability is 0.55. The female number in sampling is 0.55*40=22.
  